---
title: "UHURU Data Set Visualization"
author: "Charles"
date: "2023-02-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The Working directory inside this Rmarkdown r chunk is the following:
```{r}
getwd()
```

Note: remember that working directories in a project and in an R chunk are not always the same

## 1. Describing the working data set

UHURU data set

## 2. Reading the data set

we are reading the dataset into tsv (tab separated values).
```{r eval =TRUE}
read.csv(file = "../data-raw/ACACIA_DREPANOLOBIUM_SURVEY.txt", sep = "\t")
```

Assign the data to a variable so we can wirk with it
```{r eval =TRUE}
acacia <- read.csv(file = "../data-raw/ACACIA_DREPANOLOBIUM_SURVEY.txt", sep = "\t")
```

```{r}
head(acacia)
str(acacia)


numbers <- 1:10
numbers
numbers[3:6]


acacia
```


###Accessing elements of a data frame

It is similar to what we do for vectors, but there are two dimensions to a data frame

```{r}
acacia[8]
acacia$SURVEY
acacia[, "SURVEY"]
acacia$HEIGHT
as.numeric(acacia$HEIGHT)
```

### Replacing numbers in a data frame



```{r}
numbers
```
We identified that the HEIGHT should be numeric
```{r}
acacia$HEIGHT <- as.numeric(acacia$HEIGHT)
acacia$HEIGHT
```
```{r}
acacia <- read.csv(file = "../data-raw/ACACIA_DREPANOLOBIUM_SURVEY.txt", 
                   sep = "\t",
                   na.strings = "dead")
is.numeric(acacia$HEIGHT) 
head(acacia)
```






```{r}
plot(x = acacia$HEIGHT, y = acacia$AXIS1, main = "Acacia height vs Axis")
legend(x = 1, y = 5, legend = "Acacia Trees")
```
With GGplot, we create layers

```{r}
library(ggplot2)

ggplot(data = acacia, mapping = aes(x = HEIGHT, y = AXIS1, color = TREATMENT)) +
  geom_point(size = 5) +
  labs(x = "Tree height", y = "Tree axis 1")
```

# THE DEAL WITH THIS EXPERIMENT:
How does excluding certain classes of animals change tree height in an area?

# MEASURED VARIABLES:
Variables measured include plot number, height, rainfall, circumference, flowers, buds, fruit, and ants.

# WHAT HAPPENED TO THE OPEN PLOTS?:
Open plots tended to have low tree heights, maybe because of all the grazing that happened

